Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition.

### 2019-03-16 ACII_2019_paper/ACII_2019_paper.Rmd
summary

---
title: The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions. 
affiliation:
  ## use one only of the following
  # author-columnar: true         ## one column per author
  institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields

  institution:
    - name: Dublin City University
      location: Dublin, Ireland
      email: damien.dupre@dcu.ie
      mark: 1
      author:
        - name: Damien Dupré
    - name: Grenoble Alpes University
      location: Grenoble, France
      mark: 2
      author:
        - name: Anna Tcherkassof
          email: anna.tcherkassof@univ-grenoble-alpes.fr
abstract: |
  The abstract goes here.
  On multiple lines eventually.
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
bibliography: mybibfile.bib
output: rticles::ieee_article
---

@@

## The link between emotion and facial expression thought the behaviorist approach
Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state. This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired [@ekman1992argument]. According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49]. Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them.

## The link between emotion and facial expression thought the social constructivist approach
Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist]. For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12]. Following this assumption, faces are used as tools to display signals in social interactions [@crivelli2018facial]. These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].

>>>>

---
title: Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition. 
affiliation:
  ## use one only of the following
  # author-columnar: true         ## one column per author
  institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields

  institution:
    - name: Dublin City University
      location: Dublin, Ireland
      email: damien.dupre@dcu.ie
      mark: 1
      author:
        - name: Damien Dupré
    - name: Grenoble Alpes University
      location: Grenoble, France
      mark: 2
      author:
        - name: Anna Tcherkassof
          email: anna.tcherkassof@univ-grenoble-alpes.fr
abstract: |
  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the link between subjective feelings and facial expressions is still a subject of debate. On one hand the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions. On the other hand the constructivist approach conceives emotions are socially constructed and facial expression as social messages that are not related to emotions. In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of participants recruited to perform an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier. The results show a low accuracy of human observers and of the automatic classifier to infer from the facial expression the subjective feeling of the participants recorded. Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion socially constructed and facial expression as display of social messages appears to be more likely. Then, the way to infer emotional and mental state based on facial expressions should be questioned.
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
bibliography: mybibfile.bib
output: rticles::ieee_article
---

@@

## The link between emotion and facial expression thought the behaviorist approach
Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state. This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama]. According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49]. Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them.

## The link between emotion and facial expression thought the social constructivist approach
Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist, @barrett2017emotions]. For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12]. Following this assumption, faces are used as tools to display signals in social interactions [@crivelli2018facial]. These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].

### 2019-03-16 ACII_2019_paper/ACII_2019_paper.tex
summary

\hypersetup{
            pdftitle={The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{The relation betweem feeling and expressing emotions. A comparison
between sujective feeling and facial expression recognition of
spontaneous emotions.}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here. On multiple lines eventually.
\end{abstract}

@@

Based on the behaviorist approach initiated by Darwin in \emph{The
Expression of the Emotions in Man and Animals} (Darwin 1872), facial
expression are conceived as a genuine display of individuals inner
emotional state. This hypothesis is used as a basis for the Basic
Emotion Theory (BET) which states that a set of six or seven emotions
are universally displayed and are genetically hardwired (Ekman 1992).
According to this view, ``when emotions are aroused by perception of a
social event, a set of central commands produce patterned
emotion-specific changes in multiple systems, including {[}\ldots{}{]}
facial expressions.'' (Ekman 2007, p49). Even if this theory obtained a
popular support, it fails to explain how individuals can feel emotions
without expressing them and how individuals can express emotions without
feeling them.

@@

Detractors of the Basic Emotion Theory are perceiving emotion not as
genetically hardwired but as a learnt association between a given
situation and an appropriate response (Averill 1980). For the tenants of
the constructivist approach, emotions are ``concepts'' based on past
experiences and which are ``a collection of embodied, whole brain
representations that predict what is about to happen in the sensory
environment, what the best action is to deal with impending events, and
their consequences for allostasis'' (Barrett 2017, p12). Following this
assumption, faces are used as tools to display signals in social
interactions (Crivelli and Fridlund 2018). These signals can convey
individuals' motivations and readiness (Frijda and Tcherkassof 1997) or
social messages (Fridlund and Rosenberg 1995).

@@

\leavevmode\hypertarget{ref-barrett2017theory}{}%
Barrett, Lisa Feldman. 2017. ``The Theory of Constructed Emotion: An
Active Inference Account of Interoception and Categorization.''
\emph{Social Cognitive and Affective Neuroscience} 12 (1): 1--23.

@@

>>>>

\hypersetup{
            pdftitle={Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Are facial expressions the genuine display of individuals' subjective
feeling? A comparison with human and automatic recognition.}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
While it has been taken for granted in the development of several
automatic facial expression recognition tools, the question of the link
between subjective feelings and facial expressions is still a subject of
debate. On one hand the behaviorist approach conceives emotions as
genetically hardwired and therefore being genuinely displayed through
facial expressions. On the other hand the constructivist approach
conceives emotions are socially constructed and facial expression as
social messages that are not related to emotions. In order to evaluate
the link between the subjective feeling of emotions and their
recognition based on facial expression, 232 videos of participants
recruited to perform an emotion elicitation task were annotated by 1383
human observers as well as by an automatic facial expression classifier.
The results show a low accuracy of human observers and of the automatic
classifier to infer from the facial expression the subjective feeling of
the participants recorded. Based on these results, the hypothesis of
genetically hardwired emotion genuinely display is difficult to support
whereas the idea of emotion socially constructed and facial expression
as display of social messages appears to be more likely. Then, the way
to infer emotional and mental state based on facial expressions should
be questioned.
\end{abstract}

@@

Based on the behaviorist approach initiated by Darwin in \emph{The
Expression of the Emotions in Man and Animals} (Darwin 1872), facial
expression are conceived as a genuine display of individuals inner
emotional state. This hypothesis is used as a basis for the Basic
Emotion Theory (BET) which states that a set of six or seven emotions
are universally displayed and are genetically hardwired not only in
humans (Ekman 1992) but also in different animal species (Waal 2019).
According to this view, ``when emotions are aroused by perception of a
social event, a set of central commands produce patterned
emotion-specific changes in multiple systems, including {[}\ldots{}{]}
facial expressions.'' (Ekman 2007, p49). Even if this theory obtained a
popular support, it fails to explain how individuals can feel emotions
without expressing them and how individuals can express emotions without
feeling them.

@@

Detractors of the Basic Emotion Theory are perceiving emotion not as
genetically hardwired but as a learnt association between a given
situation and an appropriate response (Averill 1980,
@barrett2017emotions). For the tenants of the constructivist approach,
emotions are ``concepts'' based on past experiences and which are ``a
collection of embodied, whole brain representations that predict what is
about to happen in the sensory environment, what the best action is to
deal with impending events, and their consequences for allostasis''
(Barrett 2017b, p12). Following this assumption, faces are used as tools
to display signals in social interactions (Crivelli and Fridlund 2018).
These signals can convey individuals' motivations and readiness (Frijda
and Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).

@@

\leavevmode\hypertarget{ref-barrett2017emotions}{}%
Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
Life of the Brain}. Houghton Mifflin Harcourt.

\leavevmode\hypertarget{ref-barrett2017theory}{}%
---------. 2017b. ``The Theory of Constructed Emotion: An Active
Inference Account of Interoception and Categorization.'' \emph{Social
Cognitive and Affective Neuroscience} 12 (1): 1--23.

@@

\leavevmode\hypertarget{ref-de2019mama}{}%
Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
They Teach Us About Ourselves}. Granta Books.
